---
title: "hw01_jrhaag2"
author: "Jeffery Haag"
date: "9/14/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Assignment

## Exercise 1 (Working with Vectors)
**(a)**
```{r}
x = c(1, 1, 2, 3, 5, 8, 13, 21, 34, 55, 89, 144)
ones = rep(1, length(x))
mean = crossprod(x,ones)/length(x)
```
**(b)**
```{r}
z = (x - mean) 
a <- z ^ 2
sdDevSquared = crossprod(a,ones)/(length(a)-1)
sdDev = sqrt(sdDevSquared)
check = sd(x)
print(sdDev)
```
## Exercise 2 (Vectors and Matrices)
**(a)**
```{r}
x0 = rep(1, 30)
x1 = (1:30) ^ 2
x2 = seq(0, 1, length.out = 30)
x3 = log(1:30, base = exp(1))
set.seed(114)
y = 5 * x0 + 1 * x1 + 6 * x2 + 3 * x3 + rnorm(n = 30, mean = 0 , sd = 1)
report = sum(y)
print(report)
```
**(b)**
```{r}
X = matrix(c(x0, x1, x2, x3), ncol = 4)
XSum = sum(X)
print(XSum)
```
**(c)**
```{r}
beta_hat = ((t(X) %*% X) ^ -1) %*% t(X) %*% y
print((beta_hat))
```
**(d)**
```{r}
subBeta_hat = beta_hat[c(2, 3, 4), 1]
subBeta_hat = subBeta_hat ^ 2
sumSub = sum(subBeta_hat)
print(sumSub)
```
**(e)**
```{r}
y_hat = X %*% beta_hat
yRep = (y - y_hat) ^ 2
yRep = sum(yRep)
print(yRep)
```
## Exercise 3 (Discrete Probability)
**(a)**
```{r}

probA =  dbinom(5, size=50, prob=0.12) 
print(probA)
```
**(b)**
```{r}

probB =  pbinom(10, size=50, prob=0.12) 
print(probB)
```
**(c)**
```{r}

probC =  pbinom(13, size=50, prob=0.12) 
print(probC)
```
**(d)**
```{r}

probD =  pbinom(9, size=50, prob=0.12) - pbinom(2, size=50, prob=0.12)
print(probD)
```
## Exercise 4 (Continuous Probability)
**(a)**
```{r}

aProb = pnorm(95, mean = 120, sd = 15)
print(aProb)
```
**(b)**
```{r}
bProb = pnorm(140, mean = 120, sd = 15, FALSE)
print(bProb)
```
**(c)**
```{r}
not = pnorm(120, mean = 120, sd = 15, FALSE) + pnorm(95, mean = 120, sd = 15, TRUE)
cProb = 1 - not
print(cProb)
```
**(d)**
```{r}
dProb = qnorm(0.05, mean = 120, sd = 15)
print(dProb)
```
**(e)**
```{r}
eProb = qnorm(0.1, mean = 120, sd = 15, FALSE)
print(eProb)
```
## Exercise 5 (Packages, Plotting)
**(a)**
```{r}
library(faraway)
```
**(b)**
```{r}
observations = NROW(diabetes)
print(observations)
variables = NCOL(diabetes)
print(variables)
```
**(c)**
```{r}
varNames = colnames(diabetes)
print(varNames)
```
**(d)**
```{r}
HDL_mean = mean(diabetes$hdl, na.rm = TRUE)

print(HDL_mean)
```
**(e)**
```{r}
chol_sd = sd(diabetes$chol, na.rm = TRUE)

print(chol_sd)
```
**(f)**
```{r}
ages = range(diabetes$age, na.rm = TRUE)

print(ages)
```
**(g)**
```{r}
female_ldl = mean(subset(diabetes, gender == "female")$hdl)

print(female_ldl)
```
**(h)**
```{r}
plot(hdl ~ weight, data = diabetes,
     xlab = "Weight",
     ylab = "HDL",
     main = "HDL vs Weight",
     col  = "blue")
```
No not really, seems to be no correleation really, possibly slight negative correllation
**(i)**
```{r}
plot(chol ~ weight, data = diabetes,
     xlab = "Weight",
     ylab = "Cholesterol",
     main = "Chol vs Weight",
     col  = "red")
```
No not really, seems to be no correleation really
## Exercise 6 (Importing Data, Plotting)

**(a)**
```{r}
library(readr)
nutrition = read_csv("nutrition.csv")
hist(nutrition$Calories,
     xlab = "Calories",
     main = "Histogram of Calories",
     border = "blue",
     col  = "white")
```
The histogram slopes downward, there is a large calorie spike around the 400 mark which is weird. Also 0 is very frequent which is weird to me because food with 0 calroies is pretty worthless for energy.
**(b)**
```{r}
nutrition = read_csv("nutrition.csv")
plot(Calories ~ Protein, data = nutrition,
     ylab = "Calories",
     xlab = "Protein",
     main = "Calories vs Protein",
     col  = "black")
```
It has a ton of data points around the low calory area. It probably has a slightly positive correlation because there are so many poitns ar 0,0 and around it.
**(c)** 
```{r}
plot(Calories ~ I(4 * Protein + 4 * Carbs + 9 * Fat + 2 * Fiber), data = nutrition,
     xlab = "Protein",
     ylab = "Calories",
     main = "Calories vs Protein",
     col  = "green")
```
It's pretty close to a straight line but not flat because i assume 4, 9, 2 are all rough estimates and not the exact calories from each nutritional item. I dont know much about it, but I would think some carbs are gonna be counted different then others as well.
